---
title: "ggplot"
author: "Brittany Shea"
date: "2022-09-29"
output: html_document
---

#Libraries
```{r}
library(tidyverse)
library(ggridges)
```

#Import Data
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

#Let's make a scatterplot
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) +
  geom_point()
```

```{r}
weather_df %>%
  drop_na() %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()
```

#Let's keep making the same plot but different
```{r}
weather_scatterplot = 
  weather_df %>%
  drop_na() %>%
  ggplot(aes(x = tmin, y = tmax)) 

weather_scatterplot +
  geom_point()
```

#Let's make more interesting plots
#helps to find out if there's anything in here that I'm interested in:
```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point() +
  geom_smooth()
```

# Let's get fancy

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) +
  geom_smooth(se = FALSE) #get rid of error bars
```

# Maybe make separate panels

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name)
```

# tmax vs tmin is boring, let's spice it up
# show seasonal trends of maximum temp; include precipitation variable
```{r}
ggplot(weather_df, aes(x = date, y = tmax, color = name)) + 
  geom_point(alpha = .3) +
  geom_smooth(se = FALSE)

ggplot(weather_df, aes(x = date, y = tmax, color = name)) + 
  geom_point(aes(size = prcp), alpha = .3) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) # make scale cleaner
```

# Some quick stuff
#using hexogram geometry
```{r}
weather_df %>%
  drop_na() %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_hex() 
```

#Histogram, violins, etc.

#Distribution of max weather across weather stations

```{r}
weather_df  %>%
  ggplot(aes(x = tmax, fill = name)) +
  geom_density(alpha = .3)
```

#Boxplots
```{r}
weather_df  %>%
  ggplot(aes(x = name, y = tmax, fill = name)) +
  geom_boxplot()
```

#Violins
#also gives you info about complete distribution; comes in handy if comparing a lot of locations, for example
```{r}
weather_df  %>%
  ggplot(aes(x = name, y = tmax, fill = name)) +
  geom_violin()
```

#OR
#density to see direct comparisons if have a lot of distributions
```{r}
weather_df %>%
  ggplot(aes(x = tmax, y = name)) +
  geom_density_ridges()
```

#Saving and embedding plots

```{r}
ggsave(
  file = "results/weather_scatterplot.pdf",
  plot = weather_scatterplot,
  width = 8, height = 5)
```

