---
title: "ggplot"
author: "Brittany Shea"
date: "2022-09-29"
output: html_document
---

#Libraries
```{r}
library(tidyverse)
```

#Import Data
```{r}
weather_df =  
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10,
    month = lubridate::floor_date(date, unit = "month")) %>%
    select(name, id, everything())
```

#`group_by()`

```{r}
weather_df %>%
  group_by(name, month)
```

#Let's group and then summarize with GROUP & SUMMARIZE
```{r}
weather_df %>%
  group_by(name,month) %>%
  summarize(
    n_obs = n()
    )
```

#Or you can use COUNT
```{r}
weather_df %>%
  count(name, month)
```

###Don't use table 

#You can have multiple summaries

```{r}
weather_df %>%
  group_by(month) %>%
  summarize(
    n_obs = n(),
    n_days = n_distinct(date))
```

#Brief aside on 2X2 tables...
```{r}
weather_df %>% 
  mutate(
    cold = case_when(
      tmax <  5 ~ "cold",
      tmax >= 5 ~ "not_cold",
      TRUE      ~ ""
  )) %>% 
  filter(name != "Waikiki_HA") %>% 
  group_by(name, cold) %>% 
  summarize(
    n_obs = n()
  ) %>% 
  pivot_wider(
    names_from = cold,
    values_from = n_obs
  )
```

#...but Janitor package does the same as 2x2 table but easier
```{r}
weather_df %>% 
  mutate(cold = case_when(
    tmax <  5 ~ "cold",
    tmax >= 5 ~ "not_cold",
    TRUE     ~ ""
  )) %>% 
  filter(name != "Waikiki_HA") %>% 
  janitor::tabyl(name, cold)
```

#General summaries

```{r}
weather_df %>%
  group_by(name, month) %>%
  summarize(
    n_obs = n(),
    mean_tmax = mean(tmax, na.rm = TRUE),
    sd_precip = sd(precip),
    median_tmax = median(tmax, na.rm = TRUE)
  )
```

```{r}
weather_df %>%
  group_by(name, month) %>%
  summarize(across(prcp:tmin, mean))
```

#This is a dataframe

```{r}
weather_df %>%
  group_by(name, month) %>%
  summarize(
    mean_tmax = mean(tmax, na.rm = TRUE)
  ) %>%
  ggplot(aes(x = month, y = mean_tmax, color = name)) +
  geom_point() + geom_line() + 
    theme(legend.position = "bottom")
```

```{r}
weather_df %>%
  group_by(name, month) %>%
  summarize(mean_tmax = mean(tmax)) %>% 
  pivot_wider(
    names_from = name,
    values_from = mean_tmax) %>% 
  knitr::kable(digits = 2) #get a knitted table for rmarkdown document
```

```{r}
weather_df %>%
  group_by(name) %>%
  mutate(
    mean_tmax = mean(tmax, na.rm = TRUE),
    centered_tmax = tmax - mean_tmax) %>% 
  ggplot(aes(x = date, y = centered_tmax, color = name)) + 
    geom_point() 
```

#Lagged observations - includes lag by 1 observations
#If need standard deviations of day to day temperature changes: just group by, mutate, & summarize 
```{r}
weather_df %>%
  group_by(name) %>%
  mutate(
    yesterday_tmax = lag(tmax),
    tmax_change = tmax - yesterday_tmax
  ) %>%
  summarize(
    sd_tmax_change = sd(tmax_change, na.rm = TRUE)
  )
```

#If you want ranks 
```{r}
weather_df %>%
  group_by(name, month) %>%
  mutate(
    tmax_rank = min_rank(tmax) #this is a ranking function
  ) %>%
  filter(tmax_rank < 4) %>%
  arrange(name, month, tmax_rank)
```

